### 2021-05-16 manuscripts/owp_fim3_2020/sections/abstract_plainTextSummary.tex
more updates

\begin{abstract}
The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide hourly forecast discharges at four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D inundation extents used for forecasting purposes to protect life and property.
The Office of Water Prediction Flood Inundation Mapping Version 3.0 `Cahaba' (OWP-FIM V3.0) was developed to provide these inundation mapping services based on the latest advancements in methodology and data sources available. 
Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, support for a two-tiered drainage density, parametrization of processing unit size, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and optimal Manning's n roughness parameter calibrated by stream order.
OWP FIM V3 was calibrated and evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP-FIM V3 instaed of NWM discharges to isolate out factors related to hydro-meterology.
The resulting maps are compared to the modeled extents from the BLE and binary contigency statistics are computed.
Across the 48 sites, Critical Success Index (CSI) and Matthew's Correlation Coefficient (MCC) were found to increase across FIM versions driven by improvements to the probability of detection (POD) via reduction in errors of omission with only a slight increase in errors of commission.
These capabalities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
%
\end{abstract}

>>>>

\begin{abstract}
The National Water Model Version 2.1 (NWM V2.1) seeks to expand upon existing systems to provide hourly forecast discharges at four time horizons (1, 3, 10, and 30 days) at over 2.7 million locations throughout the continental United States, Hawaii and Puerto Rico.
Additional modeling capabilities are required to convert these 1-D discharge forecasts into 2-D fluvial inundation extents used for forecasting purposes to protect life and property.
The Office of Water Prediction Flood Inundation Mapping Version 3.0 `Cahaba' (OWP-FIM V3.0) is an enhancemant of the Height Above Nearest Drainage (HAND) method developed to improve upon some of HAND's limitations.
Among the implemented enhancements are efficient computational performance, support for Hawaii and Puerto Rico NWM domains, support for a two-tiered drainage density, parametrization of processing unit size, enforcement of levee data, support for the latest digital elevation model and stream network data sources, advanced stream burning techniques, thalweg conditioning, and optimal Manning's n roughness parameter calibrated by stream order.
OWP FIM V3 was calibrated and evaluated at 48 Base Level Engineering (BLE) sites furnished by the Federal Emergency Management Agency (FEMA).
The modeled discharges from the BLE sections were used for the 100 and 500 year flood events as input to OWP-FIM V3 instaed of NWM discharges to isolate out factors related to hydro-meterology.
The resulting maps are compared to the modeled extents from the BLE and binary contigency statistics are computed.
Across the 48 sites, Critical Success Index (CSI) and Matthew's Correlation Coefficient (MCC) were found to increase across FIM versions driven by improvements to the probability of detection (POD) via reduction in errors of omission with only a slight increase in errors of commission.
These capabalities were implemented in an open-source package available on GitHub (\url{https://github.com/NOAA-OWP/cahaba}) for continued use and collaboration with the research community.
%
\end{abstract}

### 2021-05-16 manuscripts/owp_fim3_2020/sections/introduction.tex
more updates

The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
The term hydrofabric is used within the NWM jargon to describe the subset of hydrography comprised of geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data. 
The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale. The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages to inundation extents.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Height Above Nearest Drainage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}. 
\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
The terrain index as even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
\citeA{zheng2018river} developed methodology for determine stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.
Catchment boundary issues are noted where the interconnection of catchments are not properly accounted for \cite{zhang2018comparative,mcgehee2016modified}.
These findings suggesting improvements to HAND require advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).

@@

A study is proposed to introduce and evaluate the development of OWP FIM V3.0 which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework. 
FIM 3 was asked to meet a series of requirements to enhance its utility for continental scale FIM from the National Water Model for operational forecasting and to ensure long-term success in future developments. 
The requirements were
\begin{enumerate}
\item use easily accessible, open-source dependencies to promote free-use and open-access community development and research.
\item excellent single-core computational and memory performance with favorable, time and space complexities,
\item ability to parallelize across multiple cores and nodes within HPC or cloud compute environments,
\item preprocessing functionality within package pipeline to acquire and process input artifacts,
\item rapid inundation mapping capabilities to go from FIM hydrofabric datasets to forecast FIMs from NWM,
\item automated testing functionality to ensure modifications enhance forecast skill when compared to external sources,
\item ability to produce depth forecasts,
\item full NWM domain coverage (including Hawaii and Puerto Rico),
\item provide initial mitigation to catchment boundary issues at RnR mainstems,
\item and improved FIM forecasting skill when compared to FIM V1.0 and V2.0.
\end{enumerate} 
To meet these higher specifications for the next generation of authoritative FIM capabilities, the OWP FIM V3.0 `Cahaba' was developed. 
It meets all the specifications listed above including version on version improvement of FIM skill due to the implementation of new developments in HAND techniques. 
The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM from the NWM for operational forecast applications. 
Furthermore, it enables the contribution from the broader hydro-community as further advances are made in this developing area of research.

>>>>

The National Hydrography Dataset Plus (NHDPlus) V2.1 is the basis for the hydrofabric in the NWM due to its comprehensive use with the hydrologic communities' stakeholders \cite{mckay2012nhdplus}. 
The term hydrofabric is used within the NWM jargon to describe the subset of hydrography comprised of geospatial datasets required for hydrologic modeling including but not limited to stream networks, catchments, channel properties, and elevation data. 
The Muskingam-Cunge routing method is used within the NWM to reduce computational requirements of a continental scale model \cite{bedient2008hydrology,ponce1994variable,gochis2018wrf}.
Muskingam-Cunge routing scheme has been demonstrated by \citeA{cunge1969subject} to be equivalent to the convective-diffusive wave method without consideration to wave dampening.
As a result of high computational costs and large spatial domains, the need for high-resolution FIM at 10m or better requires additional post-processing from the principal output of the NWM which is forecast river discharges at the reach scale. The Height Above Nearest Drainage (HAND) terrain model is one such technique that can be used, along with synthetic rating curves (SRC), to convert riverine discharges to stages to inundation extents.
\subsection{Height Above Nearest Drainage}
HAND normalizes topography along the nearest drainage path and its been demonstrated to be a good proxy and indicator of a series of important environmental conditions including soil environments, landscape classes, soil gravitational potentials, geomorphologies, soil moisture, and ground water dynamics \cite{renno2008hand,nobre2011height}. 
\citeA{nobre2016hand} showed evidence for utilizing the drainage normalizing HAND dataset as a proxy for flood potential to make static flood inundation maps from known stages.
The terrain index as even gone on to provide additional utility in the observation of riverine flood inundation mapping from remote sensing especially in areas of high electromagnetic interference such as vegetated and anthropogenic areas \cite{aristizabal2020high,shastry2019using,huang2017comparison,twele2016sentinel}.
\citeA{zheng2018river} developed methodology for determine stage-discharge relationships known as synthetic rating curves (SRC) by sampling reach-averaged parameters from HAND datasets and inputting into the Manning's equation \cite{gauckler1867etudes,manning1890flow}.
This collection of methods, coupling HAND with SRCs, has been experimented and compared to other sources of FIM including engineering scale models, in-situ observation, and remote sensing based observation with solid results in large spatial scale applications \cite{godbout2019error,johnson2019integrated,garousi2019terrain,nobre2016hand,afshari2018comparison,zheng2018geoflood,teng2015rapid,teng2017flood,zhang2018comparative}.

@@

Most notably, HAND has been shown to demonstrate sensitivity to drainage network density known colloaquially as the catchment boundary problem \cite{zhang2018comparative,mcgehee2016modified,li2020evaluation,nobre2016hand}.
This sensitivity is noted at higher order streams with large flows and low Froude numbers and is manifested by a constriction in the inundation extents at the junction location.
These findings suggest improvements to HAND are required that utilize advanced computational algorithms and software to compute a FIM hydrofabric required for producing continental-scale FIM (CFIM).

A study is proposed to introduce and evaluate the development of OWP FIM V3.0 `Cahaba' which utilizes some of the latest mapping techniques along with a few novel developments in a computationally friendly framework. 
The following methods and results describe the work in more detail and demonstrates its efficacy in producing enhanced FIM from the NWM for operational forecast applications. 
Furthermore, it enables the contribution from the broader hydro-community through its use of open-source code as further advances are made in this developing area of research.
%

### 2021-05-16 manuscripts/owp_fim3_2020/sections/keypoints.tex
more updates

\begin{keypoints}
\item KP1
\item KP2
\item KP3
\end{keypoints}

>>>>

\begin{keypoints}
\item Height Above Nearest Drainage (HAND) is a terrain index that detrends elevation data to the nearest relevant drainage path and can be used for rapid generation of flood inundation maps (FIM) across large spatial scales and high spatial resolutions specifically with applications with the United States National Water Model.
\item Despite its advantages, FIM generated from HAND suffers from a number of limitations most notably its dependence on drainage density of the stream network employed.
\item The Office of Water Prediction at the National Oceanic and Atmospheric Administration presents a means of resolving stream density by reducing drainage density to a singular flow path for HAND based FIM. More recent developments in FIM based HAND are also employed in this continental scale model. 
\end{keypoints}

### 2021-05-16 manuscripts/owp_fim3_2020/sections/materials_and_methods.tex
more updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
FIM 3 is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Dependencies and Architecture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
Within the Python 3 ecosystem, many common packages are employed including but not limited RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}. 
To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in Docker image (\url{https://docs.docker.com/engine/install/}). 
A user only need to install Docker on their host machine and build the image from the provided recipe.

The FIM 3 pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. Preprocessing acquires and prepares datasets for production of the FIM 3 hydrofabric. 
Producing the FIM hydrofabric produces the datasets required to make an inundation map including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
\subsection{Datasets}
All data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for FIM 3 while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
The AHPS forecast points are required to determine the RnR downstream segments in order to define the RnR Mainstems (MS).
For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study. 
These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
The full input datasets sorted by source are listed below:
\begin{enumerate}
\item{USGS - NHDPlusHR Beta}
    \begin{enumerate}
    \item{\textit{BurnLineEvents:} flowlines used by NHDPlusHR for hydro-enforcing DEM}
    \item{\textit{Value-Added Attributes:} database of additional attributes referenced to flowlines that enhance navigation, analysis, and display}
    \item{\textit{DEM:} DEM from National Elevation Dataset (NED) in 10m (or 1/3 arc-second) spatial resolution with vertical units in centimeters \cite{gesch2002national}}
    \end{enumerate}
\item{NOAA OWP - NWM V2.1 Hydrofabric}
    \begin{enumerate}
    \item{\textit{flowlines:} stream network center lines used for NWM routing and forecasting}
    \item{\textit{reach level catchments:} surface drainage area corresponding to each river reach}
    \item{\textit{waterbodies:} includes reservoirs, lakes, and others that are modeled within the NWM}
    \end{enumerate}
\item{USACE - NLD}
    \begin{enumerate}
    \item{\textit{Levee elevations:} Top of levee elevations are gathered with in the NLD provided by USACE for hydro-enforcement}
    \end{enumerate}
\item{Land Sea Border}
\item{FEMA BLE}
    \begin{enumerate}
    \item{\textit{cross-sections:} cross-sections and associated data for associating 1\% and 0.2\% recurrence discharges with NWM reaches}
    \item{\textit{flood inundation maps:} FIMs produced by BLE at the two recurrence intervals for comparison with FIMs from OWP FIM versions}
    \end{enumerate}
\item{NOAA OWP - AHPS Forecast Points}
    \begin{enumerate}
    \item{\textit{AHPS:} Collection AHPS forecast points to associate to RnR Mainstems}
    \end{enumerate}
\end{enumerate}
\subsection{Hydro-conditioning}
The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic. 
It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.

@@

Small reaches can lead to unrealistic variances in channel geometries while too large reaches can lead to grouping too much geometry variance into one parametrized unit.

@@

@@

where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
All units are international given the 1 numerator above n.
The reach averaged method has been evaluated to ...
The reach averaged parameters are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.

The Manning's n values are tuned ...

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{RnR Mainstems Method}
%

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inundation Mapping}
%

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and Testing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
FIM 3 is a fully operational pipeline of software tools to help acquire datasets, cache hydrofabrics, produce FIMs, and evaluate results.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Dependencies and Architecture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
OWP FIM exclusively utilizes free and open source software dependencies including Python 3, GDAL, TauDEM, Geographic Resource Analysis Support System (GRASS), GNU Parallel, and MPICH \cite{python382,gdal2020,tarboton2005terrain,grass2020,tange2015gnu,amer2021mpich}.
Within the Python 3 ecosystem, many common packages are employed including but not limited to RichDEM, GeoPandas, Rasterio, Rasterstats, and Numba \cite{barnes2018richdem,jordahl2014geopandas,lam2015numba}. 
To simplify setup and enhance portability across host operating systems OWP FIM packages all dependencies up in Docker image (\url{https://docs.docker.com/engine/install/}). 
A user only need to install Docker on their host machine and build the image from the provided recipe.

The FIM 3 pipeline is discretized into key areas that a user can interact with to reproduce the results of this study. Preprocessing acquires and prepares datasets for production of the FIM 3 hydrofabric. 
Producing the FIM hydrofabric produces the datasets required to make an inundation map including the relative elevation model (REM) or HAND grid, the catchments in vector and raster form, and the hydro-table (contains synthetic rating curves and cross-walk information).
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Datasets}
\label{ssec:datasets}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Most data sources used within FIM 3 are publicly available from a variety of government sources including the USGS, NWC, Federal Emergency Management Agency (FEMA), and US Army Core of Engineers (USACE) to enhance reproducibility and collaboration among government, academia, and industry.
The National Hydrography Dataset Plus High Resolution (NHDPlusHR) Beta Version is the latest hydrography dataset used for land surface hydrologic modeling in the US \cite{moore2019user}. 
It is used in conjunction with the hydrofabric of the NWM V2.1 to help define flowlines for FIM 3 while the NWM hydrofabric is also used to define reservoirs for exclusion and catchments to cross-walk to for forecasting purposes.
For enforcing levee data into the NED DEMs, the USACE National Levee Database (NLD) is used via burning feature elevations \cite{engineers2016national}.
Additionally, FEMA Base Level Engineering (BLE) from Region 6 (parts of Texas, Oklahoma, Arkansas, Louisiana and New Mexico) 1\% (100 year) and 0.2\% (500 year) datasets are used as validation in this study \cite{fema2021base,fema2021estimated}. 
These BLE datasets are provided at the watershed scale (HUC8) utilizing best available simulations and DEMs.
The full input datasets presented by source are list in Table \ref{tab:data}.
\note[Fernando Aristizabal]{Check Brian Avant, Trevor Grout, Bradford Bates, Ryan Spies}
%

\begin{table}
\caption{Data sources, names, and descriptions.}
\label{tab:data}
\centering
\begin{tabular}{|p{2cm}|p{4cm}|p{8cm}|}
%\begin{tabular}{l c c}
\hline
Source & Name & Description \\
\hline
USGS & NHDPlusHR BurnLineEvents & Stream lines used by NHDPlus HR for hydro-enforcement \\
\hline
USGS & NHDPlusHR Value-Added Attributes & Database of additional attributes associated with the BurnLineEvents that enhance navigation, analysis, and display \\
\hline
USGS & NHDPlusHR DEM & DEM used for NHDPlus HR at 1/3 arc-second (10m) spatial resolution and vertical units in centimeters \\
\hline
NOAA-OWP & NWM Streams & Stream network center lines used by NWM for routing and forecasting \\
\hline
NOAA-OWP & NWM Catchments & Surface drainage area corresponding to each reach in the NWM \\
\hline
NOAA-OWP & NWM Waterbodies & Waterbodies considered by the NWM as resevoirs or lakes \\
\hline
USACE & NLD & Levee database of locations and elevations  \\
\hline
WHERE? & Land-Sea Border & Border of land and sea and land and Great Lakes \\
\hline
FEMA & Cross-Sections & HEC-RAS cross-sections used for modeling in BLE datasets. Include discharges for 1\% and 0.2\% recurrence interval events \\
\hline
FEMA & Flood Inundation Extents & Extents produced by FEMA BLE for 1\% and 0.2\% recurrence interval events \\
\hline
\multicolumn{2}{l}{$^{a}$Footnote text here.}
\end{tabular}
\end{table}

Areas with all the required data (from the NWM and the USGS) are labeled as the FIM domain. 
\note[Fernando Aristizabal]{For Nick Chadwich} This includes XXXX HUC4's, XXX HUC'6s, and XX HUC8's.
An enhancement of OWP FIM v2 `Cahaba' over previous FIM versions is the support for Hawaii and Puerto Rico which the NWM v2.1 will cover.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hydro-conditioning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
The DEM from the NED is subject to a series of hydro-conditioning procedures to enhance it's suitability for riverine flood inundation mapping. 
These techniques are specific for making OWP FIM and differ from the conditioning methods used by the NHDPlusHR Beta \cite{moore2019user}.
Hydro-conditioning is implemented to obtain many objectives including enforcing the location of hydrologically relevant features such as flowlines, lakes, or drainage divides whether natural or anthropogenic. 
It can also be used to simulate more accurate bathymetry which is not accounted for in the NED 10m DEM \cite{gesch2002national}.

@@

Small reaches can lead to unrealistic variances in channel geometries while oversized reaches can lead to grouping too much geometry variance into one parametrized unit.
Section \ref{sssec:synthetic_rating_curve} details the derivation of the synthetic rating curves and the dependence on channel length. 

@@

\label{sssec:synthetic_rating_curve}

@@

where Q is discharge, y indicates the stage, n is the Manning's n roughness coefficient, V is volume at stage y, S is channel slope, L is along flow length, and B is wetted bed area at stage y.
All units are international given the 1 numerator above n.
The reach averaged method has been compared to rating curves from Hydrologic Engineering Center\'s River Analysis System (HEC-RAS) and USGS gages yielding comparable results for estimating the river bottom elevation profile, channel width at given stages, and stage-discharge relationships \cite{zheng2018river}.
The reach averaged geometry parameters including number of cells, bed area, and volume are sampled from the thalweg conditioned AGREE DEM using TauDEM's catchhydrogeo utility.
Using the split reaches described in Section \ref{sssec:deriving_fim_stream_network}, the channel slope is sampled from the thalweg conditioned DEM at the end points of the reaches while the same reaches are used to the channel length.
Calibration of the Manning's n roughness coefficient is conducted by optimizing inundation extent performance to a benchmark dataset best described in Section \ref{ssec:evaluation_and_calibration}. 
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Inundation Mapping}
\label{ssec:inundation_mapping}
%
The FIM hydrofabric consisting of the relative elevations grid, catchments grid, catchment polygons, rating curve, and cross-walking data are all used to convert forecasts from the NWM into forecasts extents.
For operational situations, one would cache the FIM hydrofabric then either produce libraries of FIM for a sample of discharges or stages or also produce the FIM in near real-time (NRT).
From the cached FIM hydrofabric and design or forecast discharges including those extracted from the NWM, inundation maps can be generated at HUC 4, 6, or 8 spatial processing units in a rapid, parallel operation. 
The discharges are associated with NWM reach identifiers and cross-walked over to reach identifiers in the FIM hydrofabric.
Once associated to a mapping stream network, 
the catchments grid encoded with the reach identifiers are used to map
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation and Calibration}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{ssec:evaluation_and_calibration}
%
Evaluation of the OWP-FIM V3 Cahaba is conducted by comparison to the HEC-RAS 1D derived models produced by FEMA region 6 \cite{fema2021base,fema2021estimated}.
The maps to the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by FEMA so we used those corresponding discharges and mapping extents for evaluation.
By not using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors imputed by hydrology, forcings, hydraulic routing, etc that we would have potentially seen if we used NWM forecasted discharges.
We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates with each NWM reach. 
To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
This partially handles the influence of neighboring cross sections that could cause flow discontinuties and mass conservation issues.
While the limitions are noted, this method does best detangle the influence of exogeneous variables that we do not wish to study in this comparison.

Calibrating the Manning's n values is very important for optimizing the skill of inundation extent forecasts \cite{zheng2018river,garousi2019terrain}.
\note[Fernando Aristizabal]{For Brian Avant} In this study we propose a simplified parameterization of roughness by stream order and tune inundation performance relative to the validation data set 
%
